---
title: "Data Read in"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(scRepertoire)
library(Seurat)
library(Matrix)
library(ggplot2)
```

Read in data. Then, create single-cell (Seurat) objects from Parse and 10x data.

The Seurat package has a convenient function for reading in Parse data: `Seurat::ReadParseBio()`



Parse data as output from the split-pipe pipeline.

```{r}
# Define the base directory once at the top
base_dir <- "PATH/benchmark/parse/outs"

# Sample 234
sid <- "234"
data_path <- file.path(base_dir, sid, "WT")

mat <- ReadParseBio(data_path)
rownames(mat)[rownames(mat) == ""] <- "unknown"
cell_meta <- read.csv(file.path(data_path, "cell_metadata.csv"), row.names = 1)

seu_234 <- CreateSeuratObject(mat, names.field = 0, meta.data = cell_meta, project = paste0("Parse_", sid))
seu_234 <- RenameCells(seu_234, new.names = paste0("S", sid, "_", rownames(seu_234@meta.data)))

# Sample 236
sid <- "236"
data_path <- file.path(base_dir, sid, "WT")

mat <- ReadParseBio(data_path)
rownames(mat)[rownames(mat) == ""] <- "unknown"
cell_meta <- read.csv(file.path(data_path, "cell_metadata.csv"), row.names = 1)

seu_236 <- CreateSeuratObject(mat, names.field = 0, meta.data = cell_meta, project = paste0("Parse_", sid))
seu_236 <- RenameCells(seu_236, new.names = paste0("S", sid, "_", rownames(seu_236@meta.data)))

# Sample 237
sid <- "237"
data_path <- file.path(base_dir, sid, "WT")

mat <- ReadParseBio(data_path)
rownames(mat)[rownames(mat) == ""] <- "unknown"
cell_meta <- read.csv(file.path(data_path, "cell_metadata.csv"), row.names = 1)

seu_237 <- CreateSeuratObject(mat, names.field = 0, meta.data = cell_meta, project = paste0("Parse_", sid))
seu_237 <- RenameCells(seu_237, new.names = paste0("S", sid, "_", rownames(seu_237@meta.data)))

# Sample 238
sid <- "238"
data_path <- file.path(base_dir, sid, "WT")

mat <- ReadParseBio(data_path)
rownames(mat)[rownames(mat) == ""] <- "unknown"
cell_meta <- read.csv(file.path(data_path, "cell_metadata.csv"), row.names = 1)

seu_238 <- CreateSeuratObject(mat, names.field = 0, meta.data = cell_meta, project = paste0("Parse_", sid))
seu_238 <- RenameCells(seu_238, new.names = paste0("S", sid, "_", rownames(seu_238@meta.data)))

# Merge into 1 (Seurat) object. Cells will be combined.
seu.combined <- merge(seu_234, y = c(seu_236, seu_237, seu_238))

# Read in TCR data. 
# The barcode report files are an output of split-pipe and contain clonotype information.

# Sample 234
sid <- "234"
S234 <- read.csv(file.path(base_dir, sid, "D14/barcode_report.tsv"), sep = "\t")

# Sample 236
sid <- "236"
S236 <- read.csv(file.path(base_dir, sid, "D14/barcode_report.tsv"), sep = "\t")

# Sample 237
sid <- "237"
S237 <- read.csv(file.path(base_dir, sid, "D14/barcode_report.tsv"), sep = "\t")

# Sample 238
sid <- "238"
S238 <- read.csv(file.path(base_dir, sid, "D14/barcode_report.tsv"), sep = "\t")

parse.contig_list <- list(S234, S236, S237, S238)
# generate a contig list
parse.contig_list <- loadContigs(input = parse.contig_list,
                                 format = "ParseBio")

# combine expression will add the immune receptor info (from combineTCR) to a Seurat or SCE object
parse.TCR <- combineTCR(parse.contig_list, samples = c("S234", "S236", "S237", "S238"), filterMulti = T)

#Adding clonotypes to meta data of the merged Seurat object
seu_combined_Parse <- combineExpression(parse.TCR, seu.combined)
```

Create single-cell (Seurat) objects from 10x data as output from the cellranger pipeline. 

```{r}
# Define the 10x base directory
base_dir_10x <- "PATH/benchmark/10x"

# QEH234
sid <- "234"
data_dir <- file.path(base_dir_10x, sid)
list.files(data_dir)
expression_matrix <- Read10X(data.dir = data_dir)
seu_234 = CreateSeuratObject(counts = expression_matrix, project = paste0("TenX_", sid))
seu_234 <- RenameCells(seu_234, new.names = paste0("S", sid, "_", rownames(seu_234@meta.data)))

# QEH236
sid <- "236"
data_dir <- file.path(base_dir_10x, sid)
list.files(data_dir)
expression_matrix <- Read10X(data.dir = data_dir)
seu_236 = CreateSeuratObject(counts = expression_matrix, project = paste0("TenX_", sid))
seu_236 <- RenameCells(seu_236, new.names = paste0("S", sid, "_", rownames(seu_236@meta.data)))

# QEH237
sid <- "237"
data_dir <- file.path(base_dir_10x, sid)
list.files(data_dir)
expression_matrix <- Read10X(data.dir = data_dir)
seu_237 = CreateSeuratObject(counts = expression_matrix, project = paste0("TenX_", sid))
seu_237 <- RenameCells(seu_237, new.names = paste0("S", sid, "_", rownames(seu_237@meta.data)))

# QEH238
sid <- "238"
data_dir <- file.path(base_dir_10x, sid)
list.files(data_dir)
expression_matrix <- Read10X(data.dir = data_dir)
seu_238 = CreateSeuratObject(counts = expression_matrix, project = paste0("TenX_", sid))
seu_238 <- RenameCells(seu_238, new.names = paste0("S", sid, "_", rownames(seu_238@meta.data)))

# merge into 1 seu object
seu.combined <- merge(seu_234, y = c(seu_236, seu_237, seu_238))

# read in contig files
S234 <- read.csv(file.path(base_dir_10x, "234", "filtered_contig_annotations.csv"))
S236 <- read.csv(file.path(base_dir_10x, "236", "filtered_contig_annotations.csv"))
S237 <- read.csv(file.path(base_dir_10x, "237", "filtered_contig_annotations.csv"))
S238 <- read.csv(file.path(base_dir_10x, "238", "filtered_contig_annotations.csv"))

# Make a list of contigs
Exp.contigs <- list(S234, S236, S237, S238)

# Sort contigs into indivudal clonotypes by barcode
combined.TCRs <- scRepertoire::combineTCR(Exp.contigs, samples = c("S234", "S236", "S237", "S238"), filterMulti = T)

#Adding clonotypes to meta data of the merged Seurat object
seu_combined_TenX <- combineExpression(combined.TCRs, seu.combined)
```

Merge 10x and Parse data to one Seurat object

```{r}
# first add metadata column "tech"
seu_combined_Parse$tech <- "Parse"
seu_combined_TenX$tech <- "10x"
# one object
combined.TCR <- c(seu_combined_Parse, seu_combined_TenX)
# seu
combined.TCR_seu_filt <- merge(seu_combined_Parse, y=seu_combined_TenX_filt)
```

Extra (optional)

-JoinLayers within Seurat object. Important for some downstream analyses.

-SingleCellExperiment is an alternative data object to Seurat. Better in some ways and also allows useful interaction with a lot of bioconductor packages. Convert with ```Seurat::as.SingleCellExperiment()```. However ```scRepetoire``` is used for downstream immune repertoire analysis and conveniently works with both objects.

```{r}
# seu obj layers
seu_combined_Parse <- JoinLayers(object = seu_combined_Parse)
seu_combined_TenX <- JoinLayers(object = seu_combined_TenX)
combined.TCR_seu <- JoinLayers(object = combined.TCR_seu)
# check meta-data
head(meta_data, 2)

# convert to SCE (bioconductor)
library(SingleCellExperiment)
library(scater)
sce_combined <- as.SingleCellExperiment(x = combined.TCR_seu)
sce_combined <- scater::logNormCounts(x = sce_combined)
seu.combined2 <- as.Seurat(x = sce_combined)
```

Save data objects

```{r, eval=FALSE}
saveRDS(object = seu_combined_Parse, file = "seu_combined_Parse.rds")
saveRDS(object = seu_combined_TenX, file = "seu_combined_10x.rds")
saveRDS(object = combined.TCR_seu, file = "seu_combined_tech_benchmark.rds")
saveRDS(object = combined.TCR_seu_filt, file = "seu_combined_filt.rds")
```
